A Monte Carlo Synthetic Acceleration Method for the $SP_N$ Equations

### 2013-03-02 doc/spn/spn.tex
working on multigroup formulation

\subsection{The Neutron Transport Equation}
\label{subsec:transport_eq}
As a starting point we define the time-independent neutron transport
equation \cite{lewis_computational_1993}:
\begin{multline}
  \hat{\Omega} \cdot \vec{\nabla} \psi(\vec{r},\hat{\Omega},E) +
  \sigma(\vec{r},E) \psi(\vec{r},\hat{\Omega},E) = \\ \int dE' \int
  d\Omega' \sigma_s(\vec{r},E' \rightarrow E,\hat{\Omega}' \rightarrow
  \hat{\Omega}) \psi(\vec{r},\hat{\Omega}',E') +
  q(\vec{r},\hat{\Omega},E)\:,
  \label{eq:general_transport}
\end{multline}
with the variables defined as:
\begin{itemize}
\item $\vec{r}$ - neutron spatial position
\item $\hat{\Omega}$ - neutron streaming direction with radial
  component $\mu$ and azimuthal component $\omega$
\item $\hat{\Omega}' \cdot \hat{\Omega} = \mu_0$ is the angle of
  scattering
\item $E$ - neutron energy
\item $\psi(\vec{r},\hat{\Omega},E)$ - angular flux
\item $\sigma(\vec{r},E)$ - total interaction cross section
\item $\sigma_s(\vec{r},E' \rightarrow E,\hat{\Omega}')$ - probability
  of scattering from direction $\hat{\Omega}'$ into an angular domain
  $d\hat{\Omega}'$ about the direction $\hat{\Omega}$ and from energy
  $E'$ to an energy domain $dE'$ about energy $E$
\item $q(\vec{r},\hat{\Omega},E)$ - external source of neutrons.
\end{itemize}
For this work, it is sufficient to formulate
Eq~(\ref{eq:general_transport}) in 1-dimensonal Cartesian geometry:
\begin{multline}
  \mu \frac{\partial}{\partial x} \psi(x,\mu,E) + \sigma(x,E)
  \psi(x,\mu,E) = \\ \int dE' \int d\Omega' \sigma_s(x,E' \rightarrow
  E,\hat{\Omega}' \rightarrow \hat{\Omega})
  \psi(\vec{r},\hat{\Omega}',E') + \frac{q(x,E)}{4 \pi}\:,
  \label{eq:cart_1d_transport}
\end{multline}
where the angular component of the solution is no longer dependent on
the azimuthal direction of travel and an isotropic source of neutrons
is assumed.

@@

\subsubsection{Legendre Polynomials}
\label{subsubsec:legendre_polys}
The Legendre polynomials are an orthogonal set of functions that
are solutions to Legendre's differential equation. They have the
following form \cite{lewis_computational_1993}:
\begin{equation}
  P_l(\mu) = \frac{1}{2^l l!}\frac{d^l}{d \mu^l}(\mu^2-1)^l\:.
  \label{eq:general_legendre_poly}
\end{equation}
These functions have several nice properties including orthogonality:
\begin{equation}
  \int_{-1}^{1} P_l(\mu) P_{l'}(\mu) d\mu = \frac{1}{2l+1}\delta_{l l'}\:,
  \label{eq:legendre_orthog}
\end{equation}
a recurrence relation:
\begin{equation}
  \mu P_l(\mu) = \frac{1}{2l+1}[(l+1)P_{l+1}(\mu) + l P_{l-1}(\mu)]\:,
  \label{eq:legendre_recurrence}
\end{equation}
and an addition theorem:
\begin{equation}
  P_l(\hat{\Omega} \cdot \hat{\Omega}') = \frac{1}{2n+1}\sum_{m=-l}^l
  Y_{lm}(\hat{\Omega})Y^*_{lm}(\hat{\Omega}')\:,
  \label{eq:legendre_addition}
\end{equation}
where the functions $Y_{lm}(\hat{\Omega})$ are the spherical
harmonics. We can form the addition theorem in this way because the
spherical harmonics are in fact just harmonic multiples of the
Legendre polynomials:
\begin{equation}
  Y_{lm}(\hat{\Omega}) =
  \sqrt{\frac{(2l+1)(l-m)!}{(l+m)!}}P^m_l(\mu)e^{i \omega m}\:,
  \label{eq:spherical_harmonic}
\end{equation}
where $\omega$ is the azimuthal component of the streaming
direction. We can reduce Eq~(\ref{eq:legendre_addition}) for the
planar geometry we are studying by ignoring the azimuthal components
of the addition theorem. As shown in Eq~(\ref{eq:spherical_harmonic}),
the azimuthal dependence is given by the harmonic component, $e^{i
  \omega m}$, and therefore we choose to ignore all terms in
Eq.~(\ref{eq:legendre_addition}) where $m \neq 0$. This gives:
\begin{equation}
  P_l(\hat{\Omega} \cdot \hat{\Omega}') = \frac{1}{2n+1}
  Y_{l0}(\hat{\Omega})Y^*_{l0}(\hat{\Omega}')\:.
  \label{eq:legendre_addition_2}
\end{equation}
Per Eq~(\ref{eq:spherical_harmonic}) we have:
\begin{equation}
  Y_{l0}(\hat{\Omega}) = \sqrt{2l+1}P_l^0(\mu)\:,
  \label{eq:harmonic_0}
\end{equation}
where $P_l^0(\mu) = P_l(\mu)$ is the $0^{th}$ associated Legendre
function. Finally, we can reduce the addition theorem from
Eq~(\ref{eq:legendre_addition_2}) with Eq~(\ref{eq:harmonic_0}) to a
simple product for planar geometry:
\begin{equation}
  P_l(\hat{\Omega} \cdot \hat{\Omega}') = P_l(\mu)P_l(\mu')\:.
  \label{eq:legendre_addition_3}
\end{equation}

@@

    \frac{16}{35}u_4\\

@@

\subsubsection{Boundary Conditions for the $SP_N$ Equations}
\label{subsubsec:spn_boundary_conditions}
We approach the formulation of the boundary conditions in much the
same way as we did for the $P_N$ equations with both reflecting and
Marshak conditions possible. To begin, we perform the expansion in
Eq~(\ref{eq:marshak_expanded}) for the left side of the planar system,
this time for $N=7$ to correspond to our $SP_7$ system. For $i=1$:
\begin{multline}
  \int_0^1 \mu \Bigg[ \phi_0 + 3\phi_1\mu +
    \frac{5}{2}\phi_2(3\mu^2-1) + \frac{7}{2}\phi_3(5\mu^3-3\mu)
    +\\ \frac{1}{8}(63\mu^5-70\mu^3+15\mu) \Bigg] d\mu = \int_0^1 \mu
  \phi_b d\mu\:,
  \label{eq:spn_bnd_p1}
\end{multline}
for $i=3$,
\begin{multline}
  \int_0^1 \frac{1}{2}(5\mu^3-3\mu) \Bigg[ \phi_0 + 3\phi_1\mu +
    \frac{5}{2}\phi_2(3\mu^2-1) + \frac{7}{2}\phi_3(5\mu^3-3\mu) +\\
    \frac{1}{8}(63\mu^5-70\mu^3+15\mu) \Bigg] d\mu = \int_0^1
  \frac{1}{2}(5\mu^3-3\mu) \phi_b d\mu\:,
  \label{eq:spn_bnd_p3}
\end{multline}
for $i=5$:
\begin{multline}
  \int_0^1 \frac{1}{8}(63\mu^5-70\mu^3+15\mu) \Bigg[ \phi_0 + 3\phi_1\mu +
    \frac{5}{2}\phi_2(3\mu^2-1) + \frac{7}{2}\phi_3(5\mu^3-3\mu) +\\
    \frac{1}{8}(63\mu^5-70\mu^3+15\mu) \Bigg] d\mu = \int_0^1
  \frac{1}{8}(63\mu^5-70\mu^3+15\mu) \phi_b d\mu\:,
  \label{eq:spn_bnd_p5}
\end{multline}
and for $i=7$:
\begin{multline}
  \int_0^1 \frac{1}{16}(429\mu^7-693\mu^5+315\mu^3+35\mu) \Bigg[
    \phi_0 + 3\phi_1\mu + \frac{5}{2}\phi_2(3\mu^2-1) +
    \frac{7}{2}\phi_3(5\mu^3-3\mu) \\+
    \frac{1}{8}(63\mu^5-70\mu^3+15\mu) +
    \frac{1}{16}(429\mu^7-693\mu^5+315\mu^3+35\mu) \Bigg] d\mu =\\
  \int_0^1 \frac{1}{16}(429\mu^7-693\mu^5+315\mu^3+35\mu) \phi_b
  d\mu\:.
  \label{eq:spn_bnd_p7}
\end{multline}
Carrying out the simple integrations yields the following system of
equations:
\begin{subequations}
  \begin{gather}
    \frac{1}{2}\phi_0 + \phi_1 + \frac{5}{8}\phi_2 -
    \frac{3}{16}\phi_4 + \frac{13}{128}\phi_6 =
    \frac{1}{2}\phi_{b}\\ -\frac{1}{8}\phi_0 + \frac{5}{8}\phi_2 +
    \phi_3 + \frac{81}{128}\phi_4 - \frac{13}{64}\phi_6 =
    -\frac{1}{8}\phi_{b}\\ \frac{1}{16}\phi_0 - \frac{25}{128}\phi_2 +
    \frac{81}{128}\phi_4 + \phi_5 + \frac{325}{512}\phi_6 =
    \frac{1}{16}\phi_{b}\\ -\frac{5}{128}\phi_0 + \frac{7}{64}\phi_2 -
    \frac{105}{512}\phi_4 + \frac{325}{512}\phi_6 + \phi_7 =
    -\frac{5}{128}\phi_{b}\:,\\
  \end{gather}
  \label{eq:spn_bnd_integrated}
\end{subequations}
where $\phi_b$ is again an isotropic source prescribed on the planar
boundary. For the odd moment in each boundary equation, we insert
Eq~(\ref{eq:odd_moments}) to remove them, leaving only the even moments:

\subsection{Formulation of the Multigroup $SP_N$ Equations}
\label{subsec:mg_spn_equations}

>>>>

\subsection{The Neutron Transport Equation}
\label{subsec:transport_eq}
As a starting point we define the time-independent neutron transport
equation \cite{lewis_computational_1993}:
\begin{multline}
  \hat{\Omega} \cdot \vec{\nabla} \psi(\vec{r},\hat{\Omega},E) +
  \sigma(\vec{r},E) \psi(\vec{r},\hat{\Omega},E) = \\ \int \int
  \sigma_s(\vec{r},E' \rightarrow E,\hat{\Omega}' \rightarrow
  \hat{\Omega}) \psi(\vec{r},\hat{\Omega}',E') d\Omega' dE' +
  q(\vec{r},\hat{\Omega},E)\:,
  \label{eq:general_transport}
\end{multline}
with the variables defined as:
\begin{itemize}
\item $\vec{r}$ - neutron spatial position
\item $\hat{\Omega}$ - neutron streaming direction with radial
  component $\mu$ and azimuthal component $\omega$
\item $\hat{\Omega}' \cdot \hat{\Omega} = \mu_0$ is the angle of
  scattering
\item $E$ - neutron energy
\item $\psi(\vec{r},\hat{\Omega},E)$ - angular flux
\item $\sigma(\vec{r},E)$ - total interaction cross section
\item $\sigma_s(\vec{r},E' \rightarrow E,\hat{\Omega}')$ - probability
  of scattering from direction $\hat{\Omega}'$ into an angular domain
  $d\hat{\Omega}'$ about the direction $\hat{\Omega}$ and from energy
  $E'$ to an energy domain $dE'$ about energy $E$
\item $q(\vec{r},\hat{\Omega},E)$ - external source of neutrons.
\end{itemize}
For this work, it is sufficient to formulate
Eq~(\ref{eq:general_transport}) in 1-dimensonal Cartesian geometry:
\begin{multline}
  \mu \frac{\partial}{\partial x} \psi(x,\mu,E) + \sigma(x,E)
  \psi(x,\mu,E) = \\ \int \int \sigma_s(x,E') \rightarrow
  E,\hat{\Omega}' \rightarrow \hat{\Omega}) \psi(x,\hat{\Omega}',E')
  d\Omega' dE' + \frac{q(x,E)}{4 \pi}\:,
  \label{eq:cart_1d_transport}
\end{multline}
where the angular component of the solution is no longer dependent on
the azimuthal direction of travel and an isotropic source of neutrons
is assumed.

@@

\subsubsection{Legendre Polynomials}
\label{subsubsec:legendre_polys}
The Legendre polynomials are an orthogonal set of functions that
are solutions to Legendre's differential equation. They have the
following form \cite{lewis_computational_1993}:
\begin{equation}
  P_l(\mu) = \frac{1}{2^l l!}\frac{d^l}{d \mu^l}(\mu^2-1)^l\:.
  \label{eq:general_legendre_poly}
\end{equation}
These functions have several useful properties including
orthogonality:
\begin{equation}
  \int_{-1}^{1} P_l(\mu) P_{l'}(\mu) d\mu = \frac{1}{2l+1}\delta_{l l'}\:,
  \label{eq:legendre_orthog}
\end{equation}
a recurrence relation:
\begin{equation}
  \mu P_l(\mu) = \frac{1}{2l+1}[(l+1)P_{l+1}(\mu) + l P_{l-1}(\mu)]\:,
  \label{eq:legendre_recurrence}
\end{equation}
and an addition theorem:
\begin{equation}
  P_l(\hat{\Omega} \cdot \hat{\Omega}') = \frac{1}{2n+1}\sum_{m=-l}^l
  Y_{lm}(\hat{\Omega})Y^*_{lm}(\hat{\Omega}')\:,
  \label{eq:legendre_addition}
\end{equation}
where the functions $Y_{lm}(\hat{\Omega})$ are the spherical
harmonics. We can form the addition theorem in this way because the
spherical harmonics are in fact just harmonic multiples of the
Legendre polynomials:
\begin{equation}
  Y_{lm}(\hat{\Omega}) =
  \sqrt{\frac{(2l+1)(l-m)!}{(l+m)!}}P^m_l(\mu)e^{i \omega m}\:,
  \label{eq:spherical_harmonic}
\end{equation}
where $\omega$ is the azimuthal component of the streaming
direction. We can reduce Eq~(\ref{eq:legendre_addition}) for the
planar geometry we are studying by ignoring the azimuthal components
of the addition theorem. As shown in Eq~(\ref{eq:spherical_harmonic}),
the azimuthal dependence is given by the harmonic component, $e^{i
  \omega m}$, and therefore we choose to ignore all terms in
Eq.~(\ref{eq:legendre_addition}) where $m \neq 0$. This gives:
\begin{equation}
  P_l(\hat{\Omega} \cdot \hat{\Omega}') = \frac{1}{2n+1}
  Y_{l0}(\hat{\Omega})Y^*_{l0}(\hat{\Omega}')\:.
  \label{eq:legendre_addition_2}
\end{equation}
Per Eq~(\ref{eq:spherical_harmonic}) we have:
\begin{equation}
  Y_{l0}(\hat{\Omega}) = \sqrt{2l+1}P_l^0(\mu)\:,
  \label{eq:harmonic_0}
\end{equation}
where $P_l^0(\mu) = P_l(\mu)$ is the $0^{th}$ associated Legendre
function. Finally, we can reduce the addition theorem from
Eq~(\ref{eq:legendre_addition_2}) with Eq~(\ref{eq:harmonic_0}) to a
simple product for planar geometry:
\begin{equation}
  P_l(\hat{\Omega} \cdot \hat{\Omega}') = P_l(\mu)P_l(\mu')\:.
  \label{eq:legendre_addition_3}
\end{equation}

@@

    \frac{16}{35}u_4 \\

@@

\subsubsection{Boundary Conditions for the $SP_N$ Equations}
\label{subsubsec:spn_boundary_conditions}
We approach the formulation of the boundary conditions in much the
same way as we did for the $P_N$ equations with both reflecting and
Marshak conditions possible. To begin, we perform the expansion in
Eq~(\ref{eq:marshak_expanded}) for the left side of the planar system,
this time for $N=7$ to correspond to our $SP_7$ system. For $i=1$:
  \int_0^1 \mu \Bigg[ \phi_0 + 3\phi_1\mu +
    \frac{5}{2}\phi_2(3\mu^2-1) + \frac{7}{2}\phi_3(5\mu^3-3\mu)
    +\\ \frac{1}{8}(63\mu^5-70\mu^3+15\mu) \Bigg] d\mu = \int_0^1 \mu
  \phi_b d\mu\:,
  \label{eq:spn_bnd_p1}
for $i=3$,
  \int_0^1 \frac{1}{2}(5\mu^3-3\mu) \Bigg[ \phi_0 + 3\phi_1\mu +
    \frac{5}{2}\phi_2(3\mu^2-1) + \frac{7}{2}\phi_3(5\mu^3-3\mu) +\\
    \frac{1}{8}(63\mu^5-70\mu^3+15\mu) \Bigg] d\mu = \int_0^1
  \frac{1}{2}(5\mu^3-3\mu) \phi_b d\mu\:,
  \label{eq:spn_bnd_p3}
for $i=5$:
  \int_0^1 \frac{1}{8}(63\mu^5-70\mu^3+15\mu) \Bigg[ \phi_0 + 3\phi_1\mu +
    \frac{5}{2}\phi_2(3\mu^2-1) + \frac{7}{2}\phi_3(5\mu^3-3\mu) +\\
    \frac{1}{8}(63\mu^5-70\mu^3+15\mu) \Bigg] d\mu = \int_0^1
  \frac{1}{8}(63\mu^5-70\mu^3+15\mu) \phi_b d\mu\:,
  \label{eq:spn_bnd_p5}
and for $i=7$:
  \int_0^1 \frac{1}{16}(429\mu^7-693\mu^5+315\mu^3+35\mu) \Bigg[
    \phi_0 + 3\phi_1\mu + \frac{5}{2}\phi_2(3\mu^2-1) +
    \frac{7}{2}\phi_3(5\mu^3-3\mu) \\+
    \frac{1}{8}(63\mu^5-70\mu^3+15\mu) +
    \frac{1}{16}(429\mu^7-693\mu^5+315\mu^3+35\mu) \Bigg] d\mu =\\
  \int_0^1 \frac{1}{16}(429\mu^7-693\mu^5+315\mu^3+35\mu) \phi_b
  d\mu\:.
  \label{eq:spn_bnd_p7}
Carrying out the simple integrations yields the following system of
equations:
    \frac{1}{2}\phi_0 + \phi_1 + \frac{5}{8}\phi_2 -
    \frac{3}{16}\phi_4 + \frac{13}{128}\phi_6 =
    \frac{1}{2}\phi_{b}\\ -\frac{1}{8}\phi_0 + \frac{5}{8}\phi_2 +
    \phi_3 + \frac{81}{128}\phi_4 - \frac{13}{64}\phi_6 =
    -\frac{1}{8}\phi_{b}\\ \frac{1}{16}\phi_0 - \frac{25}{128}\phi_2 +
    \frac{81}{128}\phi_4 + \phi_5 + \frac{325}{512}\phi_6 =
    \frac{1}{16}\phi_{b}\\ -\frac{5}{128}\phi_0 + \frac{7}{64}\phi_2 -
    \frac{105}{512}\phi_4 + \frac{325}{512}\phi_6 + \phi_7 =
    -\frac{5}{128}\phi_{b}\:,
  \label{eq:spn_bnd_integrated}
where $\phi_b$ is again an isotropic source prescribed on the planar
boundary. For the odd moment in each boundary equation, we insert
Eq~(\ref{eq:odd_moments}) to remove them, leaving only the even
moments and a set of differential equations:
    \frac{1}{2}\phi_0 + \frac{1}{3\Sigma_1}\frac{\partial}{\partial
      x}(\phi_0+2\phi_2) + \frac{5}{8}\phi_2 - \frac{3}{16}\phi_4 +
    \frac{13}{128}\phi_6 = \frac{1}{2}\phi_{b}\\ -\frac{1}{8}\phi_0 +
    \frac{5}{8}\phi_2 + \frac{1}{7\Sigma_3}\frac{\partial}{\partial
      x}(3\phi_2 + 4\phi_4) + \frac{81}{128}\phi_4 -
    \frac{13}{64}\phi_6 = -\frac{1}{8}\phi_{b}\\ \frac{1}{16}\phi_0 -
    \frac{25}{128}\phi_2 + \frac{81}{128}\phi_4 +
    \frac{1}{11\Sigma_5}\frac{\partial}{\partial x}(5\phi_4 + 6\phi_6)
    + \frac{325}{512}\phi_6 =
    \frac{1}{16}\phi_{b}\\ -\frac{5}{128}\phi_0 + \frac{7}{64}\phi_2 -
    \frac{105}{512}\phi_4 + \frac{325}{512}\phi_6 +
    \frac{1}{15\Sigma_7}\frac{\partial}{\partial x}(7\phi_6)
    = -\frac{5}{128}\phi_{b}\:.
  \label{eq:spn_bnd_subs}
We can again make the substitution of variables given by
Eqs~(\ref{eq:spn7_subs}) and (\ref{eq:spn7_subs_inverse}) for
consitency with the equations defined on the domain. In addition, we
apply the $SP_N$ approximation to the derivatives by assuming they are
instead multidimensional gradients on the boundary such that
$\frac{\partial}{\partial x} \rightarrow
\hat{\mathbf{n}}\cdot\nabla$. Doing this gives:
    \frac{1}{2}\Big(u_1 - \frac{2}{3}u_2 + \frac{8}{15}u_3 -
    \frac{16}{35}u_4\Big) +
    \frac{1}{3\Sigma_1}\hat{\mathbf{n}}\cdot\nabla\Big(\Big(u_1 -
    \frac{2}{3}u_2 + \frac{8}{15}u_3 -
    \frac{16}{35}u_4\Big)+\\2\Big(\frac{1}{3}u_2 - \frac{4}{15}u_3 +
    \frac{8}{35}u_4\Big)\Big) + \frac{5}{8}\Big(\frac{1}{3}u_2 -
    \frac{4}{15}u_3 + \frac{8}{35}u_4\Big)
    -\\ \frac{3}{16}\Big(\frac{1}{5}u_3 - \frac{6}{35}u_4\Big) +
    \frac{13}{128}\Big(\frac{1}{7}u_4\Big) = \frac{1}{2}\phi_{b}\:,
    -\frac{1}{8}\Big(u_1 - \frac{2}{3}u_2 + \frac{8}{15}u_3 -
    \frac{16}{35}u_4\Big) + \frac{5}{8}\Big(\frac{1}{3}u_2 -
    \frac{4}{15}u_3 + \frac{8}{35}u_4\Big)
    +\\ \frac{1}{7\Sigma_3}\hat{\mathbf{n}}\cdot\nabla\Big(3\Big(\frac{1}{3}u_2
    - \frac{4}{15}u_3 + \frac{8}{35}u_4\Big) + 4\Big(\frac{1}{5}u_3 -
    \frac{6}{35}u_4\Big)\Big) + \frac{81}{128}\Big(\frac{1}{5}u_3
    -\\ \frac{6}{35}u_4\Big) - \frac{13}{64}\Big(\frac{1}{7}u_4\Big) =
    -\frac{1}{8}\phi_{b}\:,
    \frac{1}{16}\Big(u_1 - \frac{2}{3}u_2 + \frac{8}{15}u_3 -
    \frac{16}{35}u_4\Big) - \frac{25}{128}\Big(\frac{1}{3}u_2 -
    \frac{4}{15}u_3 +\\ \frac{8}{35}u_4\Big) +
    \frac{81}{128}\Big(\frac{1}{5}u_3 - \frac{6}{35}u_4\Big) +
    \frac{1}{11\Sigma_5}\hat{\mathbf{n}}\cdot\nabla\Big(5\Big(\frac{1}{5}u_3
    -\\ \frac{6}{35}u_4\Big) + 6\Big(\frac{1}{7}u_4\Big)\Big) +
    \frac{325}{512}\Big(\frac{1}{7}u_4\Big) = \frac{1}{16}\phi_{b}\:,
    -\frac{5}{128}\Big(u_1 - \frac{2}{3}u_2 + \frac{8}{15}u_3 -
    \frac{16}{35}u_4\Big) + \frac{7}{64}\Big(\frac{1}{3}u_2 -
    \frac{4}{15}u_3 + \frac{8}{35}u_4\Big)
    -\\ \frac{105}{512}\Big(\frac{1}{5}u_3 - \frac{6}{35}u_4\Big) +
    \frac{325}{512}\Big(\frac{1}{7}u_4\Big)
    +\\ \frac{1}{15\Sigma_7}\hat{\mathbf{n}}\cdot\nabla\Big(7\Big(\frac{1}{7}u_4\Big)\Big)
    = -\frac{5}{128}\phi_{b}\:.
By rearranging the system such that we have a single gradient operator
in each equation, we again have a matrix system:
  \hat{\mathbf{n}} \cdot \mathbf{D} \nabla \mathbf{u} + \mathbf{B}
  \mathbf{u} = \mathbf{s}\:,
  \label{eq:spn_bnd_matrix}
where $\mathbf{D}$ and $\mathbf{u}$ are defined as before and:
  \mathbf{s} = \Big(\frac{1}{2}\phi_b\ -\frac{1}{8}\phi_b\
  \frac{1}{16}\phi_b\ -\frac{5}{128}\phi_b \Big)^T\:,
  \label{eq:spn_bnd_source}
is the source vector on the boundary and
%% I grabbed this matrix from Tom's tech note again.
  \mathbf{B} = \begin{bmatrix}
    \frac{1}{2} &
    -\frac{1}{8} &
    \frac{1}{16} &
    -\frac{5}{128} \\
    -\frac{1}{8} &
    \frac{7}{24} &
    -\frac{41}{384} &
    \frac{1}{16} \\
    \frac{1}{16} &
    -\frac{41}{384} &
    \frac{407}{1920} &
    -\frac{233}{2560} \\
    -\frac{5}{128} &
    \frac{1}{16} &
    -\frac{233}{2560} &
    \frac{3023}{17920}
  \end{bmatrix}\:,
  \label{eq:B}
is a dense matrix of coefficients. Again, as pointed out by Evans, the
$SP_1$ approximation reduces this boundary condition to the standard
Marshak diffusion boundary condition. For reflecting boundary
conditions, we use the same procedure as the $P_N$ equations where the
odd-moments are zero such that Eq~(\ref{eq:reflecting_condition}) is
true. From setting Eq~(\ref{eq:odd_moments}) to zero for odd $\phi_n$
and again substituting Eq~(\ref{eq:spn7_subs}) we immediately find
that:
  \nabla \mathbf{u} = 0
  \label{eq:spn_reflecting}
for reflecting $SP_N$ boundaries, providing enough equations to close
the system.

\subsection{Formulation of the Multigroup $SP_N$ Equations}
\label{subsec:mg_spn_equations}
Up to this point, we have formulated the $P_N$ and subsequently the
$SP_N$ equations for a single neutron energy. To expand these
equations for multiple energies, we start by stating the multigroup
neutron transport equation for a single dimension for planar geometry:
\begin{multline}
  \mu \frac{\partial}{\partial x} \psi^g(x,\mu) + \sigma^g(x)
  \psi^g(x,\mu) = \\ \sum_{g'=0}^{G} \int
  \sigma_s^{gg'}(x,\hat{\Omega}' \rightarrow \hat{\Omega})
  \psi^{g'}(x,\hat{\Omega}') d\Omega' + \frac{q^g(x)}{4 \pi}\:,
  \label{eq:cart_1d_multigroup}
\end{multline}
where $g$ denotes the group index of $0$ to $G$ groups where $G=N_g-1$
and the integration of the scattering emission term of energy has been
replaced by a discrete summation. For scattering, $\sigma_s^{gg'}$
provides the probability of scattering at a particular angle from
group $g$ to $g'$. The result is an equation nearly identical in form
to Eq~(\ref{eq:cart_1d_transport}) where now instead of forming the
$P_N$ and $SP_N$ equations for a single energy group, we form them for
each of the energy groups with group coupling occuring through the
scattering term. The multigroup $P_N$ equations are then:
\begin{equation}
   \frac{1}{2n+1} \frac{\partial}{\partial x}\Big[ (n+1) \phi^g_{n+1}
     + n \phi^g_{n-1} \Big] +
   \sum_{g'}(\sigma^g\delta_{gg'}-\sigma^{gg'}_{sn}) \phi^g_n =
   q\delta_{n0} \:,
  \label{eq:multigroup_pn_equations}
\end{equation}
for $n = 0,1,\dotsc,N$ where the flux and scattering moments are
defined in a group. We observe that a $N_g \times N_g$ scattering
matrix is formed:
\begin{equation}
  \mathbf{\Sigma}_n =
  \sum_{g'}(\sigma^g\delta_{gg'}-\sigma^{gg'}_{sn})\:,
  \label{eq:scattering_matrix}
\end{equation}
or expanded:
%% Also grabbed this matrix from Tom's tech note.
\begin{equation}
  \mathbf{\Sigma}_n =
  \begin{bmatrix}
    (\sigma^0-\sigma_{sn}^{00}) & -\sigma_{sn}^{01} & \dots &
    -\sigma_{sn}^{0G} \\ &&&\\ -\sigma_{sn}^{10} &
    (\sigma^1-\sigma_{sn}^{11}) & \dots & -\sigma_{sn}^{1G}
    \\ &&&\\ \vdots & \vdots & \ddots & \vdots
    \\ &&&\\ -\sigma_{sn}^{G0} & -\sigma_{sn}^{G1} & \dots &
    (\sigma^G-\sigma_{sn}^{GG})
  \end{bmatrix}\:.
\end{equation}
It is also useful to combine the group flux moments and sources into a
single vector for more compact notation:
\begin{equation}
  \mathbf{\Phi_n} = (\phi^0_n\ \phi^1_n\ \cdots \phi^G_n )^T\:, 
  \label{eq:group_flux_vector}
\end{equation}
\begin{equation}
  \mathbf{q} = (q^0\ q^1\ \cdots q^G )^T\:.
  \label{eq:group_source_vector}
\end{equation}
Next, we apply the $SP_N$ approximation to
Eq~(\ref{eq:multigroup_pn_equations}) in identical fashion to the
monoenergetic case. This gives:
\begin{multline}
  -\nabla \cdot \Bigg[\frac{n}{2n+1}\mathbf{\Sigma_{n-1}}^{-1} \nabla
    \Big(\frac{n-1}{2n-1} \mathbf{\Phi_{n-2}} +
    \frac{n}{2n-1}\mathbf{\Phi_n} \Big) \\+
    \frac{n+1}{2n+1}\mathbf{\Sigma_{n+1}}^{-1} \nabla
    \Big(\frac{n+1}{2n+3}\mathbf{\Phi_n} +
    \frac{n+2}{2n+3}\mathbf{\Phi_{n+2}}\Big) \Bigg] \\+
  \mathbf{\Sigma_n} \mathbf{\Phi_n} = \mathbf{q}
  \delta_{n0}\ \ \ \ \ \ \ \ \ n = 0,2,4,\cdots,N\:.
  \label{eq:multigroup_spn_equations}
\end{multline}
This adds more complexity than the monoenergetic formulation in that
all unknowns in this group of equations are now vector quantities and
scattering relationships are contained in matrices rather than a
scalar quantity.
